---
output: word_document
---
## Demonstration of the Central Limit Theorem

<br/>
<br/>

###Overview
We compare distribution of means 1000 samples (n=40) of exponentially distributed (lambda = 0.2) random numbers with the theoretical expectation based on the Central Limit Theorem

<br/>
<br/>

###Simulation
1000 iterations: generating a random sample (sample size = 40) 
of exponentially distributed data points with lambda = 0.2

```{r}
## mns is the array of the sample means - 1000 in total
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(40,0.2)))
```

<br/>
<br/>

### Sample Mean vs. Theoretical Mean
Theoretically, we expect the sample means to be centred on the real mean of the exponential distribution (1/lambda  = 5).
This is close approximated by the simulated data:

```{r}
mean(mns)
```

We plot a histogram of the sample means - to show a 'bell' type (Gaussian) distribution, centred on x = 5
```{r}
library(ggplot2)
df_mns <- data.frame(mns)

g <- ggplot(df_mns, aes(x=mns)) + geom_histogram(binwidth=.2, colour="black", fill="white")
g <- g + geom_vline(xintercept = 5)
g <- g + xlab("means of samples of 40 random exponential values each")
g <- g + ylab("count: total = 1000")
g

h <- ggplot(df_mns, aes(x=mns)) + geom_density(alpha=.2, fill="#FF6666") + geom_vline(xintercept=mean(mns))
h <- h + xlab("means of samples of 40 random exponential values each")
h <- h + ylab("Distribution Density")
h
```
<br/>
<br/>

### Sample Variance vs. Theoretical variance

Theoretically, we expect the distribution of the sample means to have variance = sigma^2 / n = 25 /40 = 0.625
This is closely approximated by the actual variance of the sample means:
```{r}
var(mns)
```
<br/>
<br/>


###Comparing with the Standard Normal Distribution
Standardising the distribution of the sample means
```{r}
nl_mns <- (mns-5)/(5/sqrt(40))
df_nl_mns <- data.frame(nl_mns)
```

We plot density curve of the standardised sample means, and compare with some known quantiles and cumulative densities of the Standard Normal distribution

```{r}
j <- ggplot(df_nl_mns, aes(x=nl_mns)) + geom_density(alpha=.2, fill="#FF6666") + geom_vline(xintercept=c(-1.96*0.7905694, -0.7905694, mean(nl_mns), 0.7905694, 1.96*0.7905694), linetype="dashed")
j <- j + xlab("means of samples of 40 random exponential values each")
j <- j + ylab("Distribution Density")
j
```

The above exploratory graph shows that the standardised distribution 
of the 1000 sample means is Gaussian, centred on zero, and its bell-curve
distribution resembles the Standard Normal 
<br/>
<br/>
Demonstrating the probability distribution of sample means within population mean +- 1.0*sd
(theoretical expectation: 0.68*1000 = 680)
```{r}
s1_nl_mns <- nl_mns[(nl_mns <= 1) & (nl_mns >= -1)]
length(s1_nl_mns)
```

<br/>
<br/>
Demonstrating the probability distribution of sample means within population mean +- 1.96*sd
(theoretical expectation: 0.95*1000 = 950)
```{r}
s2_nl_mns <- nl_mns[(nl_mns <= 1.96) & (nl_mns >= -1.96)]
length(s2_nl_mns)
```


